---
title: "Final Data Analysis Project"
date: "See Parts for Write-Up due Dates"
output:
  pdf_document: default
  html_document:
    df_print: paged
    fig_caption: TRUE
header-includes:  
  \usepackage{float} 
  \floatplacement{figure}{H}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


For this project you will take the role of a consultant hired by an Art historian to explore what drove prices of paintings in 18th century Paris.  They have provided you with auction price data from 1764-1780 on the sales (seller/buyer), painter, and other characteristics of paintings. 

## About the Data Analysis Project

The art historian would like to know what factors drove prices of painting, which paintings might be overvalued and which are undervalued.   It is up to you to decide what methods you want to use (frequentist or Bayesian or a combination) to answer these questions, and implement them to help to identify undervalued and overvalued paintings, as well as which features and possible interactions are at play.

You will have three data sets: a subset for training, a subset for testing, and a third subset for validation. You will be asked to do data exploration and build your model (or models) initially using only the training data. Then, you will test your model on the testing data, and finally validate using the validation data. We are challenging you to keep your analysis experience realistic, and in a realistic scenario you would not have access to all three of these data sets at once.  You will be able to see on our scoreboard how well your team is doing based on its predictive performance on the testing data.  After your project is turned in you will see the final score on the validation set.   

All members of the team should contribute equally and may be asked to answer any questions about the analysis at the final presentation.

*For your analysis create a new Rmd named "project-I.Rmd" for part I
and update accordingly rather than editing this.  Your write up should not have any of the instructions, for example. Figures should be labeled appropriately and report numbers using significant digits.  This file may be updated so do not edit this document.*

## Code:

In your write up your code should be hidden (`echo = FALSE`) so that your document is neat and easy to read. However your document should include all your code such that if I re-knit your Rmd file I should be able to obtain the results you presented.   If there is any code that you wish to highlight you may included it, but it should contribute significantly to your write up that should be directed to the art historian.

see Due dates in Sakai/Calendar for submissions

### Read in Training Data

To get started read in the training data:

```{r read-data, echo=TRUE}
load("paintings_train.Rdata")
load("paintings_test.Rdata")
```

The Code Book is in the file `paris_paintings.md` provides more information about the data.

## Part I: Simple Model 

```{r packages}
suppressMessages(library(GGally))
suppressMessages(library(dplyr))
suppressMessages(library(mice))
suppressMessages(library(caret))
suppressMessages(library(knitr))
```

### EDA

Using EDA and any numerical summaries get to know the data -  identify what you might consider the 10 best variables for predicting `logprice` using scatterplots with other variables represented using colors or symbols, scatterplot matrices or conditioning plots.


```{r}
# collinearity
table(paintings_train$Interm, paintings_train$type_intermed)
table(paintings_train$mat, paintings_train$materialCat)
```

Therefore, we can drop `Interm` and `materialCat` due to perfect collinearity.

```{r, echo = TRUE}
# missing randomly or not
missing <- lm(paintings_train$logprice ~ paintings_train$winningbiddertype=="")
summary(missing)
```

This indicates that `winningbiddertype` is not missing purely randomly and that there is a strong case for selective non-response. Similarly, `endbuyer`, `type_intermed`, `material` and `mat` are not missing purely randomly either. So we impute "n/a" into `winningbiddertype`, `endbuyer`, `type_intermed`, `Shape`, `material` and `mat` to replace "" and "-" as a new category.

We can drop `price` and `count` based on their meanings, with the former being another type of our response and the latter being completely useless since it is the same for all observations.

Besides, `sale`, `lot`, `subject`, `author` and `material` have too many levels, and thus they are not selected.

In addition, we have alrealy built the variables that contains the useful information of `authorstandard` and `winningbidder`, so we can drop them as well. Similarly, we can drop `Surface_Rect` and `Surface_Rnd`.

```{r clean data, warning = FALSE}
paintings_train_new <- paintings_train
# replace blank space and NA of categorical variables with n/a
paintings_train_new[, c(4, 6:9, 12:22, 25, 27:28, 30:32, 34:59)] <- as.data.frame(sapply(paintings_train_new[, c(4, 6:9, 12:22, 25, 27:28, 30:32, 34:59)], function(x) ifelse(x == "" | x == "-", "n/a", x)))
# clean position
paintings_train_new <- paintings_train_new[-which(paintings_train_new$position > 1),]
# clean Shape
paintings_train_new[which(paintings_train_new$Shape == "ronde"), 28] <- "round"
paintings_train_new[which(paintings_train_new$Shape == "ovale"), 28] <- "oval"
# choose variables
paintings_train_new <- paintings_train_new %>% 
  select(-winningbidder, -authorstandard, -sale, -lot, -price, -count, -subject, -author, -Interm, -Surface_Rect, -Surface_Rnd, -material, -materialCat)
# delete repeating data
paintings_train_new <- unique(paintings_train_new)
# imputate missing
set.seed(103)
imputed <- mice(paintings_train_new, m = 5)
paintings_train_new <- mice::complete(imputed)
# clean Surface
paintings_train_new <- paintings_train_new[-which(paintings_train_new$Surface == 0),]
```

```{r, fig.height = 24, fig.width = 16, fig.align = "center"}
# scatterplot
par(mfrow = c(6, 4))
for(i in c(1:7, 9:25)){
  plot(paintings_train_new[, i], paintings_train_new[, 8], xlab = colnames(paintings_train_new)[i], ylab = "logprice", cex.axis = 2, cex.lab = 2)
}
```

```{r, fig.height = 24, fig.width = 16, fig.align = "center", fig.cap = "Plots of predictors versus logprice"}
# scatterplot
par(mfrow = c(6, 4))
for(i in 26:46){
  plot(paintings_train_new[, i], paintings_train_new[, 8], xlab = colnames(paintings_train_new)[i], ylab = "logprice", cex.axis = 2, cex.lab = 2)
}
```

Since the points of `Surface` are clustered at the beginning of x axis, we can do log transformation to it to see its relationship with `logprice`.

```{r, fig.height = 4, fig.width = 4, fig.align = "center", fig.cap = "Plots of log Surface versus logprice"}
plot(log(paintings_train_new[, 15] + 1), paintings_train_new[, 8], xlab = paste("log(", colnames(paintings_train_new)[15], ")", sep = ""), ylab = "logprice")
```

```{r, cache = TRUE, fig.cap = "Variable Importance Measures in Random Forests"}
# log transformation to Surface
paintings_train_new$Surface <- log(paintings_train_new$Surface)
colnames(paintings_train_new)[18] <- "log_Surface"
```


### Build your first model

In the first model predict the auction price `price` using the transformation `logprice` using at least 10 and up to 20 predictors and any interactions to build a model using linear regression. You may use stepwise model selection to simplify the model using AIC and/or BIC.  For reference, we will fit the null model to initialize the leaderboard, but replace model1 with your recommended model.

### version 1

From the result of random forest, we can choose $12$ variables. (drop `diff_origin` since it's perfectly collinear with `origin_author` and `origin_cat`; drop `winningbiddertype` and `mat` since there are too many categories in them)

```{r, echo = TRUE, cache = TRUE}
# full model
model_full <- lm(logprice ~ (position + dealer + year + origin_author + origin_cat + authorstyle + endbuyer + type_intermed + log_Surface + finished + lrgfont + nfigures)^2, data = paintings_train_new)

paintings_train_new2=paintings_train_new
paintings_train_new2$nfigures[which(paintings_train_new2$nfigures > 0)] <- 1
paintings_train_new2$nfigures <- as.factor(paintings_train_new2$nfigures)
#final model
 model_test <- lm(logprice ~ (dealer + year + origin_author + origin_cat + authorstyle + endbuyer + type_intermed + log_Surface + finished + lrgfont + Shape + winningbiddertype)^2, data = paintings_train_new2)
 model_bic2 <- step(model_test, k = log(nobs(model_test)), direction = "both", trace = F)


```

```{r, warning = FALSE}
# 10-fold cross validation
set.seed(3)
train(logprice ~ dealer + year + origin_author + endbuyer + type_intermed + log_Surface + finished + lrgfont + year:endbuyer, 
      data = paintings_train_new, 
      method = "lm",
      trControl = trainControl(method = "cv", number = 10))
```

10-fold cross validation RMSE is $1.243051$.

```{r, warning = FALSE, fig.height = 8, fig.width = 8, fig.align = "center", fig.cap = "Diagnostic Plots"}
model1 <- model_bic2
# diagnostic plots
par(mfrow = c(2, 2))
plot(model1)
```

Save predictions and intervals. 

```{r predict-model1, echo = FALSE, warning = FALSE}
# test data
paintings_test_new <- paintings_test
# replace blank space and NA of categorical variables with n/a
paintings_test_new[, c(4, 6:9, 12:22, 25, 27:28, 30:32, 34:59)] <- as.data.frame(sapply(paintings_test_new[, c(4, 6:9, 12:22, 25, 27:28, 30:32, 34:59)], function(x) ifelse(paste(x) == "" | paste(x) == "-", "n/a", paste(x))))
# clean position
paintings_test_new[which(paintings_test_new$position > 1), 3] <- paintings_test_new[which(paintings_test_new$position > 1), 3]/100
# clean Shape
paintings_test_new[which(paintings_test_new$Shape == "ronde"), 28] <- "round"
paintings_test_new[which(paintings_test_new$Shape == "ovale"), 28] <- "oval"
# choose variables
paintings_test_new <- paintings_test_new %>% 
  select(-winningbidder, -authorstandard, -sale, -lot, -price, -count, -subject, -author, -Interm, -Surface_Rect, -Surface_Rnd, -material, -materialCat)
# clean Surface
paintings_test_new[which(paintings_test_new$Surface == 0), 18] <- NA
# imputate missing
set.seed(103)
imputed <- mice(paintings_test_new, m = 5)
paintings_test_new <- mice::complete(imputed)
# log transformation to Surface
paintings_test_new$Surface <- log(paintings_test_new$Surface)
colnames(paintings_test_new)[18] <- "log_Surface"
```


### Part I Write up *Last day to submit is Dec 7 by 5; accepted until Dec 6 (5 points off if late)*

Once you are satisfied with your model, provide a write up of your data analysis project in a new Rmd file/pdf file: `Part-I-Writeup.Rmd` by copying over salient parts of your R notebook. The written assignment consists of five parts:

1. Introduction: Summary of problem and objectives (5 points)

2. Exploratory data analysis (10 points): must include three correctly labeled graphs and an explanation that highlight the most important features that went into your model building.

3. Development and assessment of an initial model (10 points)

* Initial model: must include a summary table and an explanation/discussion for variable selection and overall amount of variation explained. 

* Model selection: must include a discussion

* Residual: must include residual plot(s) and a discussion.  

* Variables: must include table of coefficients and CI



```{r}
summary(model1)
anova(model1)
```
For our initial model, we first decide to incorporate all the important predictors in EDA, then add some extra predictors by the following reason.

1. origin_cat: without knowning the paintings before the auction, the buyers can only get to know the origin of paintings through the dealer's classification. For the well investigated painting, the dealer will classify the painting correctly. For unknown painting, neither dealer and buyer know the origin. So majority of the buyers will know the origin through the dealer. So we think origin_cat would be helpful for our model beside origin_author. And we use an anova test to check that.
```{r}
ano_data=paintings_train_new%>%select(logprice,origin_cat)
summary(aov(logprice~origin_cat,data=ano_data))
```
In here, we set the logprice to multiple groups according to the origin_cat, and use anova to compare if their group means are same or dfferent. Since the p_value is smaller than 0.05. we can state that the means are differnet across groups. So we decide to use origin_author


2. finished: we believe that whether a painting is finished or not will affect the the price of the sale.
And the plot of the finished can also prove our thinking. So we choose this predictor based on both commen sense and plot analysis.

Secondly, we put the choosed main predictors and all their interactions into a full model. Then we use BIC to
choose the important predictors and interactios for us.

Finally, in the initial model, we have 8 main predictors and 1 interactions. Roughly 60% variation of dependent variables are explained by this model. By looking at the anova table of the model, all of the variables are significant at the 5% level, which indicate the variables in the model are reasonable.

```{r, warning = FALSE, fig.height = 8, fig.width = 8, fig.align = "center", fig.cap = "Diagnostic Plots"}
par(mfrow = c(2, 2))
plot(model1)
```
In the Residual vs Fitted plot, there is only a slight curve at the begining of the 0 line, which is not a serious problem. Almost all points are randomly distributed around 0, indicating no violation for the linearity assumption.

The Normal Q-Q plot indicates nearly perfect poins distribution. Almost all residuals follow a normal distribution.

In the Scale-Location plot, the red line suggest that there is a small pattern for the resuduals. The residuals will increase first and then decrease. But the points are very well randomly distributed around the 0 line, So the violation of constant variance is not large enough to be very concerning. 

In the Residuals vs Leverage plot, there are neither actually influential nor potentially influential points.
no point exceed the 3 standard residuals limit, which indicate no outlier.

Generally, the diagnostic plots tell us that the linear model we get fit the train data very well. 

```{r warning=FALSE}
coef=summary(model1)$coefficients
kable(data.frame(estimate=coef[,1],
                 CI_Low=coef[,1]-1.96*coef[,2],
                 CI_Up=coef[,1]+1.96*coef[,2]))
```
In the table above, we can see part of the variables have high estimates compared to others.
It may indicate the importance of the variables or the potential problems exist in the linear model.
There also exist several variables whose condifence inteval contain 0. These variables may either have positive or negative effects on the price. we will use more complicated model in the next part to improve the performance of the model.


4. Summary and Conclusions (10 points)

What is the (median) price for the "baseline" category if there are categorical or dummy variables in the model (add CI's)?  (be sure to include units!) Highlight important findings and potential limitations of your model.  Does it appear that interactions are important?  What are the most important variables and/or interactions?  Provide interprations of how the most important variables influence the (median) price giving a range (CI).  Correct interpretation of coefficients for the log model desirable for full points.

Provide recommendations for the art historian about features or combination of features to look for to find the most valuable paintings.

_Points will be deducted for code chunks that should not be included, etc._

*Upload write up  to Sakai any time before Dec 7th*

The baseline price is $e^{-150}$ livres, whcih is approximately 0 livres. But this value is not meaningful, as it represents the price of a painting without dealer, not sale in any year, and no buyer,etc.

According to the coefficient table we get above, predictor year and endbuyer have huge impact on the price sale.For the year, althouth its coefficient is not large compared to others, the big numeric value itself will have impact on the price. Beside the year, for the dummy variable, endbuyer is another important predictor that affect the price most.

The most two important variables are year and endbuyer. And the only interaction we have is the interaction between year and endbuyer.So it's natural to say the interaction is also improtant.

Our model also has limitations. we choose the main predictors by ourselves so we may ignore some important predictors. In the initial model, the predictor year, endbuyer, and year:endbuyer looks a little over important compard to all other variables, Which means in certain degree we can just predict the price by using 3 variables. it's questionalble for such a large data set. And we only use the linear model to fit the data, resulting a few large coefficents and standard deviations. Furthermore, the big estimates value make us hard to interpreate the model to the art historian.So we may use nonlinear model to shrink the coefficient in the following part. There may even exist some more complicated relationship in the data such as polynomial. So we still need to explore that.

For every one year after the previous year, we expect price of the painting will be $e^{0.09}$ times higher, and we are 95% confident that the fluction is between $e^{-0.07}$ to $e^{0.24}$, which is from 0.93 to 1.27.

Given all other conditions unchanged(eg:same dealer, same year,same origin,etc.), we expect the price of painting will be $e^{-91}$ times higher if the buyer is a collector. And we are 95% confident that the price fluction will be between $e^{-375}$ and $e^{193}$ times higher.

Given all other conditions unchanged(eg:same dealer, same year,same origin,etc.), we expect the price of painting will be $e^{13}$ times higher if the buyer is a dealer And we are 95% confident that the price fluction will be between $e^{-272}$ and $e^{398}$ times higher. 

Given all other conditions unchanged(eg:same dealer, same year,same origin,etc.), we expect the price of painting will be $e^{-247}$ times higher if the buyer is expert organizing the sale. And we are 95% confident that the price fluction will be between $e^{-537}$ and $e^{43}$ times higher. 

Given all other conditions unchanged(eg:same dealer, same year,same origin,etc.), we expect the price of painting will be $e^{-34}$ times higher if the buyer is unknown person. And we are 95% confident that the price fluction will be between $e^{-318}$ and $e^{250}$ times higher. 

Given all other conditions unchanged(eg:same dealer, same year,same origin,etc.), we expect the price of painting will be $e^{-40}$ times higher if the buyer is person without information. And we are 95% confident that the price fluction will be between $e^{-329}$ and $e^{248}$ times higher. 

So we sugget the art hostorians that the painting bought by dealer with larger year will have a high value.




###  Evaluation on test data for Part I

Once your write up is submitted, your models will be evaluated on the following criteria based on predictions  on the test data (20 points): 

* Bias:  Average (Yhat-Y)  positive values indicate the model tends to overestimate price (on average) while negative values indicate the model tends to underestimate price.

* Maximum Deviation:  Max |Y-Yhat| -  identifies the worst prediction  made in the validation data set.

* Mean Absolute Deviation:  Average |Y-Yhat| - the average error (regardless of sign).

* Root Mean Square Error: Sqrt Average (Y-Yhat)^2

* Coverage:  Average( lwr < Y < upr) 

In order to have a passing wercker badge, your file for predictions needs to be the same length as the test data, with three columns:  fitted values, lower CI and upper CI values in that order with names, *fit*, *lwr*, and *upr* respectively such as in the code chunk below. 

Save predictions and intervals.  
```{r predict-model-final, echo=FALSE, include=FALSE}
# change model1 or update as needed
predictions <- as.data.frame(
  exp(predict(model1, newdata = paintings_test_new, 
              interval = "pred")))
save(predictions, file="predict-test.Rdata")
```

You will be able to see your scores on the score board.  They will be initialized by a prediction based on the mean in the training data.